Simulating the Linux page cache in distributed applications with the WRENCH library

### 2020-12-09 paper_io_sim.tex
added comment
edited results and conclusion

        The Linux Page Cache plays an important role in
        the reduction of filesystem data transfer costs. With the page cache, previously
        read data can be reread directly from memory and written data can be written to
        memory before being asynchronously flushed to disk, resulting in improved I/O performance
        on slower storage devices. The performance benefits that can be obtained from using the page
        cache are determined by many factors including the total amount of memory,
        the amount of data being written (i.e. dirty data), and the percentage of available memory available for
        written data. All these factors are important when determining the impact of data transfers on
        application performance, particularly in data intensive applications.

        Characterizing the performance of applications directly on computing infrastructure
        has various shortcomings, which include, limited and shared resources, operational
        costs, unstable environments, and the difficulties in producing reproducible results.
        Simulations address these concerns by providing abstractions for the performance of computer
        hardware, such as CPU, network and storage. Consequently, simulations provide and
        cost-effective, fast, easy and reproducible  way evaluate application performance
        on different infrastructures without having to execute the code.

        The amount of data-intensive applications has been on the rise as a result of
        open-data and data sharing initiatives. Due to the sheer size of the data being
        processed, these applications must be executed on large scale infrastructure
        such as High Performance Computing (HPC) clusters or the cloud. Determining
        which type of infrastructure is the most cost-effective for the particular
        application can be tricky without the help of a simulator.

        Page cache simulation is crucial when simulating data intensive applications,
        which rely on cache hits and writing to memory for additional performance.
        Furthermore, simulating the page cache enables us to study important application-I/O
        bottlenecks which can arise over the course of execution, such as page cache
        starvation~\cite{zhuang2017}.
        While existing simulators can adequately characterize many different computing environments,
        they lack much of the ability to adequately describe the page cache. Some
        simulators try to simulator I/O with page cache, but fail to capture the key features such
        as dirty data and cache eviction policies~\cite{nunez2012simcan,nunez2012icancloud}. 
        Other simulators, such as~\cite{xu2018}, capture the important feature of the
        page cache, but are domain-specific.

@@

        In this paper, we present WRENCH-Cache, a simulation model of I/O with page
        cache incorporated into WRENCH~\cite{casanova2018wrench}, a workflow management system simulator based on
        the popular SimGrid framework. The goal of WRENCH-Cache is to provide users
        with an accurate representation of a workflow execution, when file system page
        cache is used, while providing a good balance between accuracy and scalability.
        WRENCH-Cache is able to simulate both writeback and writethrough
        cache which would normally be available to local and NFS-based filesystems.
        WRENCH-Cache can provide accurate simulations for large-scale parallel
        applications using the page cache.

@@

        To improve performance of disk accesses, the page cache was introduced to the Linux kernel. %It contains pages referring to physical pages on disk \cite{linuxdev3rd2010}.
            The goal of the page cache is to try ensure that most I/O operations occur entirely
            within memory space rather than synchronously to and from the storage device.
            All disk-backed data located in the page cache are partitioned into many small pages.
            When a read operation is initiated, the kernel checks whether the required pages are in memory.
            If it is the case, data is read entirely from memory (i.e. \textit{cache hit}). Otherwise,
            the necessary pages must be loaded to memory from the disk (i.e. \textit{cache miss})
            Conversely, for writing, one of three strategies can be applied: direct~I/O (or, ``no-write''),
            writethrough and writeback.
            If direct~I/O is enabled for a given file system, all I/O operations bypass the
            page cache for writes, writing data directly to disk. In contrast, write-through enables the page cache
            for file system writes, but the data is written synchronously from memory to disk. Writeback extends write-through
            by enabling asynchronous writes to disk, meaning that a write call will return once
            the data has been written to memory.
            The writeback strategy is considered to outperform write-through as well as 
            direct I/O as it delays disk writes to perform a bulk write at a later time.
            \cite{linuxdev3rd2010}.

            Cache eviction and flushing strategies are integral to proper page cache functioning.
            Whenever space in memory becomes limited, either as a result of application memory
            or page cache use, page cache data may be evicted. Only data that 
            has been persisted to storage (clean pages) can be flagged for eviction and removed from
            memory. Written data that has not yet been persisted to disk (dirty data) must first
            be copied (flushed) to storage prior to eviction. When sufficient memory is
            being occupied, the flushing process is synchronous. However, even when 
            there is sufficient available memory, written data will be flushed to disk
            at a predefined interval through a process known as \textit{periodical flushing}.
            Periodical flushing only flushes expired dirty pages, which remain dirty in 
            page cache longer than an expiration time configured in the kernel. 
            Different cache eviction algorithms have also been developed and proposed
            \cite{owda2014comparison}.
            The Linux kernel uses a two-list strategy to flag pages for eviction.
            The two-list strategy is based on a least recently used (LRU) policy
            and uses an active and inactive list in its implementation.
            If accessed pages are not in the page cache, they are added to the inactive list. 
            Should pages located on the inactive list be accessed, they will be moved from 
            the inactive to the active list. 
            The lists are also kept balanced by moving pages from the active list
            to the inactive list when the active list grows too large.
            Thus, the active list only contains pages which are accessed more than once 
            and not evictable, while the inactive list includes pages accessed once only, 
            or pages that have be accessed more than once but moved from the active list. 
            Both lists operate using LRU eviction policies meaning the data that has
            not be accessed recently will be moved first.

@@

            Many simulation frameworks exist to describe the behaviours of applications
            running on the various infrastructures available to users, such as personal computers,
            clouds and clusters.
            These simulators are built with resource models such CPU, network and 
            storage, using various different approaches to describe the hardware.
            For example, there is \textit{CloudSim} and \textit{iCanCloud} for packet-level network models 
            and \textit{DiskSim} the simulates storage hardware functioning.

            Simulators must choose between scalability and accuracy when describing the
            hardware.\cite{casanova2014simgrid}. 
            Fine-grained models allow simulators to achieve high accuracy, 
            but lead to scalability issues due to long simulation time 
            when simulated platforms and applications grow. 
            SimGrid is an open source simulation platform that has been developed 
            not only to balance the trade-off between accuracy and scalability, 
            but also to achieve versatility, allowing the use of the framework 
            to simulate applications in multiple domains. WRENCH, a workflow management
            system simulator based on the SimGrid framework, provides a high-level API
            for simulating workflows, thus facilitating the integration of new models.
            \todo{say more about wrench}

            Despite the fact that many sophisticated models for resources have been 
            integrated in simulators, and the page cache shows apparent impacts on I/O
            performance, the simulation model of page cache is still missing in most 
            of simulators.
            For example, SimGrid only simulates I/O operations with storage bandwidths 
            and capacity. 
            SIMCAN is a simulation framework that models page cache by storing data 
            accessed on disk in a block cache \cite{nunez2012simcan}. 
            Page cache is also modeled in the iCanCloud simulator through a component which
            manages memory accesses and cached data \cite{nunez2012icancloud}. However,
            the scalability of the iCanCloud simulator is limited by its sophisticated model.
            Ultimately, none of these simulators provide any writeback cache simulator nor
            cache eviction policys through LRU lists.
            Although cache replacement policies are applied in \cite{xu2018saving} to simulate 
            in-memory caching, their simulator is specific to energy consumption of multi-tier 
            heterogenous networks.

            In this study, we use WRENCH 
            to implement our page cache model. We chose WRENCH due to accurate yet scalable implementation, resulting
            from its SimGrid backend, as well as its high-level API.
            We have integrated our model in WRENCH and use the SimGrid API to interact with 
            simulated storage resources.
            This approach allows us to leverage existing features in WRENCH and SimGrid such as process control and synchronization of 
            resource consumption.
            We also leverage the existing SimGrid storage model with capacity 
            and bandwidth properties, which are similar to the characteristics of memory.

@@

    \begin{algorithm}\caption{Periodical flushing simulation}\label{alg:pdflush}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{in}{page cache inactive list}
                \Desc{ac}{page cache active list}
                \Desc{t}{predefined flushing time interval}
                \Desc{exp}{predefined expiration time}
                \Desc{sm}{storage simulation model}
               \EndInput
               \While{host is on}
                \State blocks = expired\_blocks(exp, in) + expired\_blocks(exp, ac) 
                \State flushing\_time = sm.write(blocks)
                \State for blk in blocks: blk.dirty = 0
                \If{flushing\_time $<$ t}
                    \State sleep(t - flushing\_time)
                \EndIf  % Else: do nothing, continue the loop
            \EndWhile
        \end{algorithmic}
    \end{algorithm}

@@

    \begin{algorithm}\caption{File chunk read simulation of IO Controller}
    \label{alg:read}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{cs}{chunk size}
                \Desc{fn}{file name}
                \Desc{fs}{file size (assumed to fit in memory)}
                \Desc{mm}{MemoryManager object}
                \Desc{sm}{storage simulation model}
               \EndInput
               \State disk\_read = min(cs, fs - mm.cached(fn)) \Comment{To be read from disk}
               \State cache\_read = cs - disk\_read \Comment{To be read from cache}
               \State required\_mem = cs + disk\_read
               \State mm.flush(cs + disk\_read - mm.free\_mem - mm.evictable, fn) 
               \State mm.evict(cs + disk\_read - mm.free\_mem, fn) 
               \If {disk\_read $>$ 0}  \Comment{Read uncached data}    
               \State sm.read(disk\_read)  
               \State mm.add\_to\_cache(disk\_read, fn)     
               \EndIf
               \If {cache\_read $>$ 0} \Comment{Read cached}
               \State mm.cache\_read(cache\_read)  
            \EndIf
            \State mm.use\_anonymous\_mem(cs)
        \end{algorithmic}
    \end{algorithm}
    As mentioned previously, our model reads and writes file chunks in a
    round-robin fashion. To read a file chunk, simulated applications send
    chunk read requests to the IO Controller which processes them using
    Algorithm~\ref{alg:read}. First, we calculate the amount of uncached
    data that needs to be read from disk, and the remaining amount is read
    from cache (line 6-7). The amount of memory required to read the chunk
    is calculated, corresponding to a copy of the chunk in anonymous memory
    and a copy of the chunk in cache (line 8). 
    If there is not enough available memory, the Memory Manager is called
    to flush dirty data (line 9). If necessary, flushing is complemented by
    eviction (line 10). Note that, when called with negative arguments, functions
    \texttt{flush} and \texttt{evict} simply return and do not do anything. Then, if there is
    uncached data, memory manager is called to read data from disk and add this 
    data to cache (line 12).
    If cached data needs to be read, the Memory Manager is called to simulate 
    a cache read  and update the corresponding data blocks accordingly (line 15).
    Finally, the memory manager is called to reduce the amount of free memory 
    corresponding to the amount of anonymous memory used by application (line 17).

    \begin{algorithm}\caption{File chunk write simulation of IO Controller}
    \label{alg:write}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{cs}{chunk size}
                \Desc{fn}{file name}
                \Desc{mm}{MemoryManager object}
               \EndInput
            \State remain\_dirty = dirty\_ratio * mm.avail\_mem - mm.dirty
            \If {remain\_dirty $>$ 0} \Comment{Write with memory bandwidth}
                \State mm.evict(min(cs, remain\_dirty) - mm.free\_mem)
                \State mem\_amt = min(cs, mm.free\_mem)
                \State mm.write\_to\_cache(fn, mem\_amt) 
            \EndIf
            \State remaining = cs - mem\_amt
            \While {remaining $>$ 0}  \Comment{Write with disk bandwidth}
                \State mm.flush(cs - mem\_amt)
                \State mm.evict(cs - mem\_amt  - mm.free\_mem) 
                \State to\_cache = min(remaining, mm.free\_mem)
                \State mm.write\_to\_cache(fn, to\_cache) 
                \State remaining = remaining - to\_cache
            \EndWhile

@@

            Our first experiment (\textbf{Exp single-threaded}) simply runs one instance of the
            application on a single cluster node, with different input file
            sizes (20~GB, 50~GB, 75~GB, 100~GB), and with all I/Os directed
            to the same local disk. The goal is to validate the simulation model 
            in absence of any concurrent disk access.

            Our second experiment (\textbf{Exp concurrent}) runs concurrent instances of the
            application on a single node, all application instances
            operating on different files of size 3~GB stored in the same
            local disk, sharing the disk bandwidth. We vary the number of
            concurrent application instances from 1 to 32 since cluster
            nodes have 32 CPU cores.

            Our third experiment (\textbf{Exp NFS}) occurs in the same configuration as the
            previous one, albeit reading and writing on a 400-GB NFS-mounted remote disk
            mounted from another compute node. As is commonly
            configured in HPC environments to avoid data loss, in our simulator, 
            NFS client and server read caches are enabled, but there is no cache 
            for writing on client side and server writes are configured with
            write-through cache instead of writeback. 
            Therefore, all the writes happen at disk bandwidth, but
            reads can benefit from cache hits. We simulate write-through by
            writing data directly to disk and adding that data to the server
            cache.

@@

        Figure~\ref{fig:single_error} shows the relative error of simulated
        read and write times for the three successive application tasks in \textbf{Exp single-threaded}. The
        page cache simulation model drastically reduces simulation errors
        for all I/O operations except the first read as it involves
        uncached data. Errors are reduced from an average of 345\%
        in the original WRENCH to 46\% in the Python prototype and 39\% in WRENCH-cache.
        Unsurprisingly, the original WRENCH simulator significantly overestimates
        read and write times, due to its lack of page cache simulation
        model.

        Especially, the simulation time errors of \textit{Read 3} with 100 GB of input
        in the simulators with page cache are just under the error in original WRENCH.
        This longer read can be explained with the amount of cached data of files
        at the beginning of the read in reality and simulators shown in
        Fig.~\ref{fig:single_cache}.
        % In this figure, \textit{file $<$i$>$} is the input of \textit{Read $<$i$>$} and
        % output of \textit{Write $<$i-1$>$}.
        With the input size of 20 GB, the amount of cached data of each files after 
        every I/O operation is similar since the page cache can keep all these files.
%             \tristan{what does this refer to?
%             For 20 GB there is still some error in WRENCH-cache, what do you mean 
%``there is no difference in the page cache''?}.
%   Dzung: Some text was lost in a merge, I added it again.
        With the input of 100 GB, the difference between real execution and WRENCH-cache
        simulator can be seen in the amount of \textit{file 3} cache after \textit{Write 2}.
        In real execution, after being written by \textit{Write 2},
        \textit{file 3} is entirely in cache.
        Thus, in \textit{Read 3}, \textit{file 3} is read with memory bandwidth.
        However, in the WRENCH-cache simulator, only a part of \textit{file 3} is
        cached after \textit{Write 2}, which means \textit{Write 2} reads a part of
        \textit{file 3} from disk.
        The reason for the difference is in our WRENCH-cache simulator,
        with 100 GB of input, while \textit{file 3} is being written,
        the page cache is saturated, eviction is then evoked to evict data
        from page cache and make space available for \textit{file 3}.
        However, because the evicted amount is less than required,
        some \textit{file 3} data is evicted after being written, make \textit{file 3}
        partially cached.
        As a result, in \textit{Read 3}, only a part of \textit{file 3} is read from cache
        with memory bandwidth, a part of it is read from disk with disk read bandwidth,
        which is many times slower, leading to longer read time.

            Memory profiling results of real pipeline execution and simulators 
            with 20GB and 100GB input files are illustrated in Fig.~\ref{fig:single_memprof}.
            Because the results with 20Gb, 50GB, 100GB shows a similar pattern,
            we only include the results with 20GB and 100GB in this paper.

            As shown in the figures, the makespans of the simulated pipeline are
            close to the makespan of the real execution with similar patterns of I/O time
            real execution and simulation.
            But more importantly, with all input sizes, similar trends can also be found
            in the lines describing memory status in simulators and the real execution.
            The similar evolutions of amount of cache used (purple lines) in reality
            and simulation show that the use of page cache is well reproduced by the simulators.
            The amount of dirty data (red lines) also follow similar trends,
            which means not only the amount dirty data, but also the flushing and
            periodical flushing mechanisms are accurately modeled in our simulators.
            This can be seen in in the increase of the amount of dirty data during the writes
            and the decrease during the reads and computations.
            According to the results achieved with all input sizes in the experiment,
            we can observe identical trends in reality and simulation with 20GB,
            50GB and 75GB input files. However, with 100GB input file,
            the third read is not correctly simulated, leading to longer simulated time
            than in real execution. This will be explained later in this section.

             Figure~\ref{fig:single_error} shows the relative error of simulated
             read and write times throughout the three application tasks. The
             page cache simulation model drastically reduces simulation errors
             for all I/O operations except the first read that deals with
             uncached data. Errors are reduced from an average of 345\%
             in WRENCH to 46\% in the Python prototype and 39\% in WRENCH-cache.
             The original WRENCH simulator significantly overestimates
             read and write times, due to its lack of page cache simulation
             model.
             Especially, the simulation time errors of \textit{Read 3} with 100 GB of input
             in the simulators with page cache are just under the error in original WRENCH.
             This longer read can be explained with the amount of cached data of files
             at the beginning of the read in reality and simulators shown in
             Fig.~\ref{fig:single_cache}.
             % In this figure, \textit{file $<$i$>$} is the input of \textit{Read $<$i$>$} and
             % output of \textit{Write $<$i-1$>$}.
             With the input size of 20 GB, the amount of cached data of each files after 
             every I/O operation is similar since the page cache can keep all these files.
%             \tristan{what does this refer to?
%             For 20 GB there is still some error in WRENCH-cache, what do you mean 
%``there is no difference in the page cache''?}.
%   Dzung: Some text was lost in a merge, I added it again.
             With the input of 100 GB, the difference between real execution and WRENCH-cache
             simulator can be seen in the amount of \textit{file 3} cache after \textit{Write 2}.
             In real execution, after being written by \textit{Write 2},
             \textit{file 3} is entirely in cache.
             Thus, in \textit{Read 3}, \textit{file 3} is read with memory bandwidth.
             However, in the WRENCH-cache simulator, only a part of \textit{file 3} is
             cached after \textit{Write 2}, which means \textit{Write 2} reads a part of
             \textit{file 3} from disk.
             The reason for the difference is in our WRENCH-cache simulator,
             with 100 GB of input, while \textit{file 3} is being written,
             the page cache is saturated, eviction is then evoked to evict data
             from page cache and make space available for \textit{file 3}.
             However, because the evicted amount is less than required,
             some \textit{file 3} data is evicted after being written, make \textit{file 3}
             partially cached.
             As a result, in \textit{Read 3}, only a part of \textit{file 3} is read from cache
             with memory bandwidth, a part of it is read from disk with disk read bandwidth,
             which is many times slower, leading to longer read time.

>>>>

        The Linux Page Cache, introduced in kernel version 2.4, plays an important role in
        the reduction of filesystem data transfer costs~\cite{pagecache}. With the page cache, previously
        read data can be reread directly from memory and written data can be written to
        memory before being asynchronously flushed to disk, resulting in improved I/O performance
        on slower storage devices. The performance benefits that can be obtained from using the page
        cache are determined by many factors including the total amount of memory,
        the amount of data being written (i.e. dirty data), and the percentage of available memory available for
        written data. All these factors are important when determining the impact of data transfers on
        application performance, particularly in data intensive applications.

        Characterizing the performance of applications directly on computing infrastructure
        has various shortcomings, which include, limited and shared resources, operational
        costs, unstable environments, and the difficulties in producing reproducible results.
        Simulations address these concerns by providing abstractions to the performance of computer
        hardware, such as CPU, network and storage. As a result, simulations provide and
        cost-effective, fast, easy and reproducible  way evaluate application performance
        on different infrastructures without having to execute the code.

        Page cache simulation is crucial when simulating data intensive applications,
        which would rely on cache hits and writing to memory for additional performance.
        While existing simulators can adequately characterize many different computing environments,
        they lack much of the ability to adequately describe the page cache. Some
        simulators try to simulator I/O with page cache, but fail to capture the key features such
        as dirty data and cache eviction policies~\cite{nunez2012simcan,nunez2012icancloud}. 
        Other simulators, such as~\cite{xu2018}, capture the important feature of the
        page cache, but are domain-specific.

@@

        In this paper, we present WRENCH-cache, a simulation model of I/O with page
        cache incorporated into WRENCH, a workflow management system simulator based of
        the popular SimGrid framework. The goal of WRENCH-cache is to provide users
        with an accurate representation of a workflow execution, when file system page
        cache is used. WRENCH-cache is able to simulate both writeback and writethrough
        cache which would normally be available to local and NFS-based filesystems.

@@

            Linux implements a disk cache called \textit{page cache}. 
            It contains pages referring to physical pages on disk \cite{linuxdev3rd2010}.
            The goal of the page cache is to avoid disk I/O by keeping data in memory 
            for later accesses, minimize I/O time since disk accesses are considerably 
            slower than memory accesses.
            When the kernel starts a read operation, it checks if required data is in page cache.
            If yes, data is read from memory. This is called a \textit{cache hit}. 
            Otherwise, data is read from disk and placed in page cache.
            For writing, page cache is implemented with three different strategies. 
            The first is \textit{direct~I/O} or \textit{no-write}, which means the page cache 
            is not involved in write operations. 
            The second is \textit{write-through}, in which data goes \textit{through} 
            page cache to disk. In this case, data is written to both page cache and disk. 
            In the third strategy, called \textit{writeback}, data are written directly 
            to disk as \textit{dirty} pages and these pages will be periodically flushed 
            to disk later by a \textit{writeback} process and then marked as not dirty. 
            The writeback strategy is considered to outperform write-through as well as 
            direct I/O as it delays disk writes and perform a bulk write later.
            \cite{linuxdev3rd2010}.

            In writeback, one of the keys is the cache eviction mechanism, which is
            triggered when available memory is insufficient to remove pages 
            from page cache. 
            The term \textit{available memory} here includes both free memory and 
            the memory used but can be reclaimed by the kernel with flushing 
            and eviction. 
            Because dirty pages are data written to page cache but yet to be 
            written to disk, dirty pages must be kept in page cache and only clean pages 
            are available for eviction.
            If the clean pages are inadequate, dirty pages are flushed to disk, 
            marked as clean and available for eviction. 
            There is another type of flushing mechanism, which is 
            \textit{periodical flushing}, called periodically in predefined intervals. 
            Periodical flushing only flushes expired dirty pages, which remain dirty in 
            page cache longer than an expiration time configured in the kernel. 
            Different cache eviction algorithms have also been developed and proposed
            \cite{owda2014comparison}.
            In Linux kernel, a \textit{two lists strategy (LRU/2)} is implemented based on 
            the least-recently-used (LRU) scheme, with an \textit{active list} and 
            an \textit{inactive list}. 
            If accessed pages are not in the page cache, they are added to the inactive list. 
            If pages are accessed when residing in the inactive list, they will be moved from 
            inactive list to active list. 
            The lists are also kept balanced by moving pages from the active list
            to the inactive list when the active list grows too large.
            Thus, the active list only contains pages which are accessed more than once 
            and not evictable, while the inactive list includes pages accessed once only, 
            or pages accessed more than once but moved from the active list. 
            The lists work in LRU manner, pages are added to the tail and removed 
            from the top.

@@

            Many simulation frameworks and applications have been developed 
            in recent years to support studies in grids, cloud, HPC, peer-to-peer 
            networks, volunteer computers.                     
            These simulators are built with models of resources such as CPU, network, 
            storage with different approaches and levels of details in different simulators. 
            The detail level of the model causes a conflict between accuracy and scalability, 
            and these simulators usually trade of one for the other 
            \cite{casanova2014simgrid}. 
            For example, some simulators are built with packet-level network models 
            such as \textit{CloudSim} and \textit{iCanCloud}, or storage model that 
            simulates the operation of storage hardware in \textit{DiskSim} 
            \cite{casanova2014simgrid}. 
            Fine-grained models allow simulators to achieve high accuracy, 
            but at the same time lead to scalability issues due to long simulation time 
            when simulated platforms and applications grow. 
            SimGrid is an open source simulation platform that has been developed 
            not only to balance the trade-off between accuracy and scalability, 
            but also to achieve versatility, which allows the use of the framework 
            to simulate applications in multiple domains \cite{casanova2014simgrid}.

            Despite the fact that many sophisticated models for resources have been 
            integrated in simulators, and the page cache shows apparent impacts on I/O
            performance, the simulation model of page cache is still missing in most 
            of simulators.
            For example, SimGrid only simulates I/O operations with storage bandwidths 
            and capacity. 
            SIMCAN is a simulation framework that models page cache by storing data 
            accessed on disk in a block cache \cite{nunez2012simcan}. 
            Page cache is also modeled in iCanCloud simulator with a component in charge 
            of managing memory accesses and cached data \cite{nunez2012icancloud}, 
            but one of the issues is scalability due to its sophisticated model.
            Ultimately, non of these simulators provide any cache write strategy, 
            cache eviction mechanism with dirty/clean data and the page cache LRU lists. 
            Cache replacement policies is applied in \cite{xu2018saving} to simulate 
            in-memory caching, but for the study in energy consumption of multi-tier 
            heterogenous network.

            In this study, we chose SimGrid to implement our model of I/O with page cache 
            due to the balancing between accuracy and scalibility, the versatility allowing it 
            to be adopted in different domains. 
            In addition, it is an actively maintained open-scource framework with 
            modular design, which enables the framework to be extended with new models. 
            The main issue users have to face when using SimGrid is everything has to be 
            implemented from scratch due to its low-level API. 
            To ease this challenge, WRENCH, another open-source library, is built on top 
            of SimGrid to abstract the low-level logic and expose a higher abstraction 
            level API \cite{casanova2018wrench}. 
            We integrate our model in WRENCH and use the SimGrid API to interact with 
            simulated storage resource.
            This approach allows us to leverage existing features in WRENCH and SimGrid.
            Firstly, we can take advantage of the process control, synchronization of 
            resource consumption by concurrent processes provided by the simulation core 
            in SimGrid. 
            Secondly, we can reuse the existing SimGrid storage model with capacity 
            and bandwidth properties, which are similar to the characteristics of memory.
            And finally, WRENCH provides a high level abstraction, which help us drop 
            the burden of implementation.

@@

    \begin{algorithm}\caption{Periodical flushing simulation in IO Controller}\label{alg:pdflush}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{in}{page cache inactive list}
                \Desc{ac}{page cache active list}
                \Desc{t}{predefined flushing time interval}
                \Desc{exp}{predefined expiration time}
                \Desc{sm}{storage simulation model}
               \EndInput
               \While{host is on}
                \State blocks = expired\_blocks(exp, in) + expired\_blocks(exp, ac) 
                \State flushing\_time = sm.write(blocks)
                \State for blk in blocks: blk.dirty = 0
                \If{flushing\_time $<$ t}
                    \State sleep(t - flushing\_time)
                \EndIf  % Else: do nothing, continue the loop
            \EndWhile
        \end{algorithmic}
    \end{algorithm}

@@

    \begin{algorithm}\caption{File chunk read simulation in IO Controller}
    \label{alg:read}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{cs}{chunk size}
                \Desc{fn}{file name}
                \Desc{fs}{file size (assumed to fit in memory)}
                \Desc{mm}{MemoryManager object}
                \Desc{sm}{storage simulation model}
               \EndInput
               \State disk\_read = min(cs, fs - mm.cached(fn)) \Comment{To be read from disk}
               \State cache\_read = cs - disk\_read \Comment{To be read from cache}
               \State required\_mem = cs + disk\_read
               \State mm.flush(cs + disk\_read - mm.free\_mem - mm.evictable, fn) 
               \State mm.evict(cs + disk\_read - mm.free\_mem, fn) 
               \If {disk\_read $>$ 0}  \Comment{Read uncached data}    
               \State sm.read(disk\_read)  
               \State mm.add\_to\_cache(disk\_read, fn)     
               \EndIf
               \If {cache\_read $>$ 0} \Comment{Read cached}
               \State mm.cache\_read(cache\_read)  
            \EndIf
            \State mm.use\_anonymous\_mem(cs)
        \end{algorithmic}
    \end{algorithm}
    As mentioned previously, our model reads and writes file chunks in a
    round-robin fashion. To read a file chunk, simulated applications send
    chunk read requests to the IO Controller which processes them using
    Algorithm~\ref{alg:read}. First, we calculate the amount of uncached
    data that needs to be read from disk, and the remaining amount is read
    from cache (line 6-7). The amount of memory required to read the chunk
    is calculated, corresponding to a copy of the chunk in anonymous memory
    and a copy of the chunk in cache (line 8). 
    If there is not enough available memory, the Memory Manager is called
    to flush dirty data (line 9). If necessary, flushing is complemented by
    eviction (line 10). Note that, when called with negative arguments, functions
    \texttt{flush} and \texttt{evict} simply return and do not do anything. Then, if there is
    uncached data, memory manager is called to read data from disk and add this 
    data to cache (line 12).
    If cached data needs to be read, the Memory Manager is called to simulate 
    a cache read  and update the corresponding data blocks accordingly (line 15).
    Finally, the memory manager is called to reduce the amount of free memory 
    corresponding to the amount of anonymous memory used by application (line 17).

    \begin{algorithm}\caption{File chunk write simulation in IO Controller}
    \label{alg:write}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{cs}{chunk size}
                \Desc{fn}{file name}
                \Desc{mm}{MemoryManager object}
               \EndInput
            \State remain\_dirty = dirty\_ratio * mm.avail\_mem - mm.dirty
            \If {remain\_dirty $>$ 0} \Comment{Write with memory bandwidth}
                \State mm.evict(min(cs, remain\_dirty) - mm.free\_mem)
                \State mem\_amt = min(cs, mm.free\_mem)
                \State mm.write\_to\_cache(fn, mem\_amt) 
            \EndIf
            \State remaining = cs - mem\_amt
            \While {remaining $>$ 0}  \Comment{Write with disk bandwidth}
                \State mm.flush(cs - mem\_amt)
                \State mm.evict(cs - mem\_amt  - mm.free\_mem) 
                \State to\_cache = min(remaining, mm.free\_mem)
                \State mm.write\_to\_cache(fn, to\_cache) 
                \State remaining = remaining - to\_cache
            \EndWhile

@@

            Our first experiment (\textit{Exp single-threaded}) simply runs one instance of the
            application on a single cluster node, with different input file
            sizes (20~GB, 50~GB, 75~GB, 100~GB), and with all I/Os directed
            to the same local disk. The goal is to validate the simulation model 
            in absence of any concurrent disk access.

            Our second experiment (\textit{Exp concurrent}) runs concurrent instances of the
            application on a single node, all application instances
            operating on different files of size 3~GB stored in the same
            local disk, sharing the disk bandwidth. We vary the number of
            concurrent application instances from 1 to 32 since cluster
            nodes have 32 CPU cores.

            Our third experiment (\textit{Exp NFS}) occurs in the same configuration as the
            previous one, albeit reading and writing on a 400-GB NFS-mounted remote disk
            mounted from another compute node. As is commonly
            configured in HPC environments to avoid data loss, in our simulator, 
            NFS client and server read caches are enabled, but there is no cache 
            for writing on client side and server writes are configured with
            write-through cache instead of writeback. 
            Therefore, all the writes happen at disk bandwidth, but
            reads can benefit from cache hits. We simulate write-through by
            writing data directly to disk and adding that data to the server
            cache.

@@

        The page cache simulation model drastically reduces I/O simulation
        errors in the three successive application tasks (Figure~\ref{fig:single_error}). The first read is not impacted 
        as it only involves uncached data. Errors are reduced from an average
        of 345\% in the original WRENCH to 46\% in the Python prototype and
        39\% in WRENCH-cache. Unsurprisingly, the original WRENCH simulator
        significantly overestimates read and write times, due to its lack
        of page cache simulation model. Results with files of 50~GB and 75~GBs
        show similar behaviors and are not reported due to space limitations.

        WRENCH simulation errors are substantially lower with 100~GB
        files than with 20~GB files, due to the fact that part of the
        100~GB file needs to be read and written to disk, the only storage
        device in WRENCH, as it does not fit in cache. Conversely,
        simulation errors of the Python prototype and WRENCH-cache are higher with
        100~GB files than with 20~GB files, due to idiosyncracies in the kernel 
        flushing and eviction strategies.

        Simulated memory profiles are highly consistent with the real ones
        (Figure~\ref{fig:single_memprof}). With 20~GB files, memory
        profiles almost exactly match the real ones, although dirty data
        seems to be flushing faster in real life than in simulation
        \tristan{any idea why?}, a behavior also observed with 100~GB files. With
        100~GB files, used memory reaches total memory during the first write,
        triggering dirty data flushing. Used memory drops to cached memory
        when application tasks release anonymous memory. Simulated cached
        memory is highly consistent with real values, except toward the end
        of Read 3 where it slightly increases in simulation but not in
        reality \tristan{why?}. As expected, dirty data remains under the
        dirty ratio. The Python prototype and WRENCH-cache exhibit nearly
        identical memory profiles, which reinforces the confidence in our
        implementation.

        The content of the simulated memory cache is also highly consistent
        with reality (Figure~\ref{fig:single_cache}). With 20~GB files, the
        simulated cache content exactly matches reality, owing to the fact
        that all files fit in page cache. With 100~GB files, a slight
        discrepancy is observed after Write 2, which explains the
        simulation error previously mentioned in Read 3. In the real execution indeed, 
        File 3 is entirely cached after Write 2, whereas in the simulated execution, part 
        of it is on disk. This is due to the fact that the simulated flushing and eviction 
        was triggered after File 3 was already written, while in reality these operations 
        happened before File 3 was written \tristan{Dzung, pls check that. Also, why doesn't it happen irl?}.

        % We conclude from this experiment that our model can accurately
        % simulate cached I/Os in a sequential context. The next experiment 
        % evaluates our simulation with concurrent I/Os.

            \begin{figure*}
            \centering
            \begin{subfigure}{\linewidth}
                \centering
                   \includegraphics[width=\linewidth]{result/single/figures/single_errors.pdf}
                   \vspace*{-0.7cm}
                   \caption{Absolute relative simulation errors}
                   \vspace*{0.5cm}
                   \label{fig:single_error}
                \end{subfigure}
            \begin{subfigure}{\linewidth}
                \centering
                %    Gray shades represent task phases (read, compute and write). 
                %    Lines represent memory usage along pipeline execution time.}
                   \includegraphics[width=\linewidth]{result/single/figures/single_memprof.pdf}
                   \vspace*{-0.7cm}
                   \caption{Memory profiles \tristan{Remove trailing whitespace for 20GB graphs (xmax should be about 160s, not 200s)}}
                   \vspace*{0.5cm}
                   \label{fig:single_memprof}
            \end{subfigure}
            \begin{subfigure}{\linewidth}
                \centering
                   \includegraphics[width=1.05\linewidth]{result/single/figures/cached_files.pdf}
                   \caption{Cache contents \emph{after} application I/O operations}
                %    \textcolor{red}{Update real results of 20GB}}
                   \label{fig:single_cache}
            \end{subfigure}
            \caption{Single-threaded results}
            \end{figure*}

        \subsection{Concurrent applications}

            \begin{figure*}
            \begin{subfigure}{\linewidth}
                \centering
                \includegraphics[width=\linewidth]{result/multi/figures/multi_local.pdf}
            \end{subfigure}
            \caption{Concurrent results with 3~GB files (averages on 5 repetitions)}
            \label{fig:multi_local}
            \end{figure*}

            The page cache model notably reduce WRENCH's simulation error
            for concurrent applications and local I/Os
            (Figure~\ref{fig:multi_local}). For reads, WRENCH-cache still
            slightly overestimates reality, due to the discrepancy between
            simulated and real read bandwidths mentioned before
            (\tristan{check it's indeed mentioned. Dzung: is it correct?}). For writes, 
            WRENCH-cache slightly underestimates reality for the same reason. The plateau 
            ending around 7 concurrent pipelines in the WRENCH-cache curve
            marks the limit beyond which the page cache is saturated with dirty data 
            and needs flushing \tristan{why is there no plateay irl?}.
